# Calibration and Sensitity

We will make use of `rechaRge` API to perform calibration and sensitivity analysis with different tools.

## Quality assessment

Following the previous example, we will need to load observations datasets:

```{r}
input_rcn_gauging <- paste0(base_url, "rcn_gauging.csv.gz") # relation between gaugins station and RCN cell IDs
input_observed_flow <- paste0(base_url, "observed_flow.csv.gz") # flow rates in mm/d
input_alpha_lyne_hollick <- paste0(base_url, "alpha_lyne_hollick.csv.gz")
```

And we also need in this case to update the settings of the **HydroBudget** model object, so that column names match with the expected ones:

```{r}
HB$rcn_gauging_columns <- list(
  rcn_id = "cell_ID",
  station_id = "gauging_stat"
)
HB$alpha_lyne_hollick_columns$station_id <- "station"
```

Then we can process the river flow observations and assess simulation quality:

```{r, message=FALSE}
quality <- rechaRge::evaluate_simulation_quality(
  HB,
  water_budget = water_budget,
  rcn_gauging = input_rcn_gauging,
  observed_flow = input_observed_flow,
  alpha_lyne_hollick = input_alpha_lyne_hollick,
  period = simul_period
)
```

The `rechaRge` package proposes an model-free implementation of the [Kling-Gupta Efficiency](https://en.wikipedia.org/wiki/Kling%E2%80%93Gupta_efficiency) algorithm, that can be used for quality evaluation. In the case of our example the quality measurements of interest are:

```{r}
list(
  KGE_qtot_cal_mean = mean(quality$simulation_metadata$KGE_qtot_cal),
  KGE_qbase_cal_mean = mean(quality$simulation_metadata$KGE_qbase_cal))
```


## Using sensitivity

The [sensitivity](https://cran.r-project.org/package=sensitivity) R package can perform various sensitivity analysis.

### Define model function

```{r}
library(rechaRge)
library(data.table)

# Preload input data
# Quiet download
options(datatable.showProgress = FALSE)
# use input example files provided by the package
base_url <- "https://github.com/gwrecharge/rechaRge-book/raw/main/examples/input/"
input_rcn <- fread(paste0(base_url, "rcn.csv.gz"))
input_climate <- fread(paste0(base_url, "climate.csv.gz"))
input_rcn_climate <- fread(paste0(base_url, "rcn_climate.csv.gz"))
input_rcn_gauging <- fread(paste0(base_url, "rcn_gauging.csv.gz"))
input_observed_flow <- fread(paste0(base_url, "observed_flow.csv.gz"))
input_alpha_lyne_hollick <- fread(paste0(base_url, "alpha_lyne_hollick.csv.gz"))
# Simulation period
simul_period <- c(2017, 2017)

hydrobudget_eval <- function(i) {
  # Calibration parameters
  HB <- rechaRge::new_hydrobugdet(
    T_m = i[1],
    # melting temperature (°C)
    C_m = i[2],
    # melting coefficient (mm/°C/d)
    TT_F = i[3],
    # Threshold temperature for soil frost (°C)
    F_T = i[4],
    # Freezing time (d)
    t_API = i[5],
    # Antecedent precipitation index time (d)
    f_runoff = i[6],
    # Runoff factor (-)
    sw_m = i[7],
    # Maximum soil water content (mm)
    f_inf = i[8] # infiltration factor (-)
  )
  # Input data specific settings
  HB$rcn_columns <- list(
    rcn_id = "cell_ID",
    RCNII = "RCNII",
    lon = "X_L93",
    lat = "Y_L93"
  )
  HB$climate_columns$climate_id <- "climate_cell"
  HB$rcn_climate_columns <- list(climate_id = "climate_cell",
                                 rcn_id = "cell_ID")
  HB$rcn_gauging_columns <- list(rcn_id = "cell_ID",
                                 station_id = "gauging_stat")
  HB$alpha_lyne_hollick_columns$station_id <- "station"

  # Simulation with the HydroBudget model
  water_budget <- rechaRge::compute_recharge(
    HB,
    rcn = input_rcn,
    climate = input_climate,
    rcn_climate = input_rcn_climate,
    period = simul_period,
    workers = 1
  )

  # Evaluate simulation quality
  result <- rechaRge::evaluate_simulation_quality(
    HB,
    water_budget = water_budget,
    rcn_gauging = input_rcn_gauging,
    observed_flow = input_observed_flow,
    alpha_lyne_hollick = input_alpha_lyne_hollick,
    period = simul_period
  )

  return(c(
    mean(result$simulation_metadata$KGE_qtot_cal),
    mean(result$simulation_metadata$KGE_qbase_cal)
  ))
}
```

### Run sensitivity analysis

```{r, message=FALSE, results='hide', cache=TRUE, fig.show='hide'}
library(sensitivity)
# Use future package to parallel
library(future.apply)

hydrobudget_sens <- function(X) {
  kge_hb <- as.matrix(t(future_apply(X, MARGIN = 1, FUN = hydrobudget_eval, future.seed = TRUE)))
  return(kge_hb)
}

# Number of variables
nvar <- 8
# Range of the parameters
binf <- c(1, 4, -20, 5, 3.05, 0.5, 160, 0.01)
bsup <- c(2.5, 6.5, -12, 30, 4.8, 0.6, 720, 0.05)

# parallel computation setting
plan(multisession, workers = 3)
#plan(sequential) # non parallel
sensitivity_results <- morris(
  model = hydrobudget_sens,
  factors = nvar,
  r = 2,
  design = list(type = "oat", levels = 5, grid.jump = 3),
  binf = binf,
  bsup = bsup)
```

### Handle sensitivity results

```{r}
# Variable of interest
mu <- apply(sensitivity_results$ee, 3, function(M){
  apply(M, 2, mean)
})
mu.star <- apply(abs(sensitivity_results$ee), 3, function(M){
  apply(M, 2, mean)
})
sigma <- apply(sensitivity_results$ee, 3, function(M){
  apply(M, 2, sd)
})
sensitivity_eval <- data.table(mu = mu, mu.star = mu.star, sigma = sigma)
```

```{r, echo=FALSE}
knitr::kable(sensitivity_eval)
```


```{r}
# Plot
plot(sensitivity_results, y_col = 2)
title(main = "y_col = 2")
```


```{r}
plot.new()
plot(sensitivity_results, y_col = 1)
title(main = "y_col = 1")
```




## Using caRamel

The [caRamel](https://cran.r-project.org/package=caRamel) R package can perform both calibration and sensitivity analysis.

### Define objective function

We will start by defining the objective function to optimize: this function will run a simulation and quality evaluation and its returned values will be used by `caRamel` to measure the quality of the injected parameters.

```{r}
make_hydrobudget_eval <- function() {
  # Preload input data
  # Quiet download
  options(datatable.showProgress = FALSE)
  # use input example files provided by the package
  base_url <- "https://github.com/gwrecharge/rechaRge-book/raw/main/examples/input/"
  input_rcn <- data.table::fread(paste0(base_url, "rcn.csv.gz"))
  input_climate <- data.table::fread(paste0(base_url, "climate.csv.gz"))
  input_rcn_climate <- data.table::fread(paste0(base_url, "rcn_climate.csv.gz"))
  input_rcn_gauging <- data.table::fread(paste0(base_url, "rcn_gauging.csv.gz"))
  input_observed_flow <- data.table::fread(paste0(base_url, "observed_flow.csv.gz"))
  input_alpha_lyne_hollick <- data.table::fread(paste0(base_url, "alpha_lyne_hollick.csv.gz"))
  # Simulation period
  simul_period <- c(2017, 2017)

  hydrobudget_eval <- function(i) {
    # Calibration parameters
    HB <- rechaRge::new_hydrobugdet(
      T_m = x[i, 1],
      # melting temperature (°C)
      C_m = x[i, 2],
      # melting coefficient (mm/°C/d)
      TT_F = x[i, 3],
      # Threshold temperature for soil frost (°C)
      F_T = x[i, 4],
      # Freezing time (d)
      t_API = x[i, 5],
      # Antecedent precipitation index time (d)
      f_runoff = x[i, 6],
      # Runoff factor (-)
      sw_m = x[i, 7],
      # Maximum soil water content (mm)
      f_inf = x[i, 8] # infiltration factor (-)
    )
    # Input data specific settings
    HB$rcn_columns <- list(
      rcn_id = "cell_ID",
      RCNII = "RCNII",
      lon = "X_L93",
      lat = "Y_L93"
    )
    HB$climate_columns$climate_id <- "climate_cell"
    HB$rcn_climate_columns <- list(climate_id = "climate_cell",
                                   rcn_id = "cell_ID")
    HB$rcn_gauging_columns <- list(rcn_id = "cell_ID",
                                   station_id = "gauging_stat")
    HB$alpha_lyne_hollick_columns$station_id <- "station"

    # Simulation with the HydroBudget model
    rechaRge::with_verbose(FALSE)
    water_budget <- rechaRge::compute_recharge(
      HB,
      rcn = input_rcn,
      climate = input_climate,
      rcn_climate = input_rcn_climate,
      period = simul_period,
      workers = 1 # do not parallelize, caRamel will do it
    )

    # Evaluate simulation quality
    quality <- rechaRge::evaluate_simulation_quality(
      HB,
      water_budget = water_budget,
      rcn_gauging = input_rcn_gauging,
      observed_flow = input_observed_flow,
      alpha_lyne_hollick = input_alpha_lyne_hollick,
      period = simul_period
    )

    return(c(
      mean(quality$simulation_metadata$KGE_qtot_cal),
      mean(quality$simulation_metadata$KGE_qbase_cal)
    ))
  }

  return(hydrobudget_eval)
}
```

### Run calibration analysis

Then perform calibration with sensitivity:

```{r, message=FALSE, results='hide', cache=TRUE, fig.show='hide'}
library(caRamel)

# Number of objectives
nobj <- 2
# Number of variables
nvar <- 8
# All the objectives are to be maximized
minmax <- c(TRUE, TRUE)
# Ranges of the parameters
bounds <- matrix(nrow = nvar, ncol = 2)
bounds[, 1] <- c(1, 4, -20, 5, 3.05, 0.5, 160, 0.01)
bounds[, 2] <- c(2.5, 6.5, -12, 30, 4.8, 0.6, 720, 0.05)

calibration_results <- caRamel(
  nobj = nobj,
  nvar = nvar,
  minmax =  minmax,
  bounds = bounds,
  func = make_hydrobudget_eval(),
  prec = matrix(0.01, nrow = 1, ncol = nobj),
  sensitivity = FALSE, # you can include sensitivity analysis
  archsize = 100,# adjust to relevant value
  popsize = 10,  # adjust to relevant value
  maxrun = 10,   # adjust to relevant value
  carallel = 1, # do parallel ...
  numcores = 2  # ... on 2 cores
)
```
### Handle calibration results

Make use of calibration results, by merging simulation outputs with front objectives, to get more readable parameters:

```{r}
output_front <- data.table::data.table(cbind(calibration_results$parameters, calibration_results$objectives))
colnames(output_front) <- c("T_m", "C_m", "TT_F", "F_T", "t_API", "f_runoff", "sw_m", "f_inf", "KGE_qtot", "KGE_qbase")
```

```{r, echo=FALSE}
knitr::kable(head(output_front))
```

Display the results with `caRamel`'s plotting feature:

```{r}
# Plot using caRamel
plot_caramel(calibration_results, objnames = c("KGE_qtot", "KGE_qbase"))
```

Scatter plot the calibration's simulations, with the resulting [Pareto front](https://en.wikipedia.org/wiki/Pareto_front):

```{r}
# Plot all using ggplot
library(ggplot2)
front <- data.table(calibration_results$objectives)
colnames(front) <- c("KGE_qtot", "KGE_qbase")
all <- data.table(calibration_results$total_pop[, 9:10])
colnames(all) <- c("KGE_qtot", "KGE_qbase")
combined_data <- rbind(all, front)
combined_data$group <- c(rep("All", nrow(all)), rep("Front", nrow(front)))
ggplot(combined_data, aes(x = KGE_qtot, y = KGE_qbase, color = group)) +
  geom_point() +
  labs(title = "Scatter plot of simulations",
       x = "KGE_qtot", y = "KGE_qbase",
       color = "Dataset") +
  theme_minimal()
```

