# Example

Load the _rechaRge_ library.

```{r}
library(rechaRge)
```

## HydroBudget model

### Input data and parameters

Load the input data for the simulation.

```{r}
base_url <- "https://github.com/gwrecharge/rechaRge-book/raw/main/examples/input/"
input_rcn <- paste0(base_url, "rcn.csv.gz") # RCN values per RCN cell ID
input_climate <- paste0(base_url, "climate.csv.gz") # precipitation total in mm/d per climate cell ID
input_rcn_climate <- paste0(base_url, "rcn_climate.csv.gz") # relation between climate and RCN cell IDs
input_rcn_gauging <- paste0(base_url, "rcn_gauging.csv.gz") # relation between gaugins station and RCN cell IDs
input_observed_flow <- paste0(base_url, "observed_flow.csv.gz") # flow rates in mm/d
input_alpha_lyne_hollick <- paste0(base_url, "alpha_lyne_hollick.csv.gz")
```

Set the **HydroBudget** model with the parameters values:

```{r}
HB <- rechaRge::new_hydrobugdet(
  T_m = 2.1, # melting temperature (°C)
  C_m = 6.2, # melting coefficient (mm/°C/d)
  TT_F = -17.6, # Threshold temperature for soil frost (°C)
  F_T = 16.4, # Freezing time (d)
  t_API = 3.9, # Antecedent precipitation index time (d)
  f_runoff = 0.63, # Runoff factor (-)
  sw_m = 431, # Maximum soil water content (mm)
  f_inf = 0.07 # infiltration factor (-)
)
```

Set the column names mappings matching the input datasets:

```{r}
HB$rcn_columns <- list(
  rcn_id = "cell_ID",
  RCNII = "RCNII",
  lon = "X_L93",
  lat = "Y_L93"
)
HB$climate_columns$climate_id <- "climate_cell"
HB$rcn_climate_columns <- list(
  climate_id = "climate_cell",
  rcn_id = "cell_ID"
)
HB$rcn_gauging_columns <- list(
  rcn_id = "cell_ID",
  station_id = "gauging_stat"
)
HB$alpha_lyne_hollick_columns$station_id <- "station"
```

Set the simulation period:

```{r}
simul_period <- c(2010, 2017)
```

### Simulation

Compute the water budget using the HydroBudget model:

```{r, message=FALSE}
water_budget <- rechaRge::compute_recharge(
  HB,
  rcn = input_rcn,
  climate = input_climate,
  rcn_climate = input_rcn_climate,
  period = simul_period
)
```

The water budget data set is per year-month in each RCN cell:

* `vi`, the vertical inflow
* `t_mean`, the mean temperature
* `runoff`, the runoff
* `pet`, the potential evapotranspiration
* `aet`, the actual evapotranspiration
* `gwr`, the ground water recharge
* `runoff_2`, the excess runoff

The head of this data set is:

```{r, echo=FALSE}
knitr::kable(head(water_budget))
```

### Quality assessment

Process the river flow observations and assess simulation quality:

```{r, message=FALSE}
result <- rechaRge::compute_simulation_quality_assessment(
  HB,
  water_budget = water_budget,
  rcn_gauging = input_rcn_gauging,
  observed_flow = input_observed_flow,
  alpha_lyne_hollick = input_alpha_lyne_hollick,
  period = simul_period
)
```

### Results handling

Save the simulation results:

```{r}
sim_dir <- file.path(tempdir(), paste0("simulation_HydroBudget_", format(Sys.time(), "%Y%m%dT%H_%M")))

# Write output files
# CSV
rechaRge::write_recharge_results(HB, water_budget, output_dir = sim_dir)
# NetCDF
rechaRge::write_recharge_results(HB, water_budget, output_dir = sim_dir, format = "nc", input_rcn = input_rcn, names = list(
  "lon" = list(
    longname = "Qc lambert NAD83 epsg32198 Est",
    unit = "m"
  ),
  "lat" = list(
    longname = "Qc lambert NAD83 epsg32198 North",
    unit = "m"
  )
))
# Rasters
rechaRge::write_recharge_rasters(
  HB,
  water_budget = water_budget,
  input_rcn = input_rcn,
  crs = "+proj=lcc +lat_1=60 +lat_2=46 +lat_0=44 +lon_0=-68.5 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs",
  output_dir = sim_dir
)


# List simulation output files
list.files(sim_dir)
```

### Data visualization

Visualize the saved raster files:

```{r, message=FALSE}
library(tidyterra)
library(terra)
library(ggplot2)
library(cowplot)
subtitle <- ifelse(simul_period[1] == simul_period[2],
  paste0("In ", simul_period[1]),
  paste0("From ", simul_period[1], " to ", simul_period[2])
)
runoff <- terra::rast(file.path(sim_dir, "interannual_runoff_NAD83.tif"))
runoffplot <- ggplot() +
  geom_spatraster(data = runoff) +
  scale_fill_viridis_c(option = "inferno") +
  labs(
    fill = "",
    title = "Runoff",
    subtitle = subtitle
  )
aet <- terra::rast(file.path(sim_dir, "interannual_aet_NAD83.tif"))
aetplot <- ggplot() +
  geom_spatraster(data = aet) +
  scale_fill_viridis_c(option = "inferno") +
  labs(
    fill = "",
    title = "Actual Evapotranspiration",
    subtitle = subtitle
  )
gwr <- terra::rast(file.path(sim_dir, "interannual_gwr_NAD83.tif"))
gwrplot <- ggplot() +
  geom_spatraster(data = gwr) +
  scale_fill_viridis_c(option = "inferno") +
  labs(
    fill = "",
    title = "Ground Water Recharge",
    subtitle = subtitle
  )
cowplot::plot_grid(runoffplot, aetplot, gwrplot)
```

```{r, include=FALSE}
# Clean up
unlink(sim_dir, recursive = TRUE)
```

