--- 
title: "rechaRge: the HydroBudget model"
author: "Emmanuel Dubois, Yannick Marcon"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: gwrecharge/rechaRge-book
description: "The book presents the rechaRge package, and how it is used with the HydroBudget model."
---


```{r, setup, include=FALSE}
knitr::opts_chunk$set(comment="", warning=FALSE, message=FALSE, cache=TRUE,
                      fig.topcaption=TRUE, fig.path = "fig/")
library(kableExtra)
```

# Welcome {-}

This book provides users with some common workflows for performing groundwater recharge modeling. This book will teach you how to use the [rechaRge](https://github.com/gwrecharge/rechaRge) R package [@R-rechaRge]. This material serves as an online companion for the manuscript “Simulation of long-term spatiotemporal variations in regional-scale groundwater recharge: contributions of a water budget approach in cold and humid climates” [@hess-25-6567-2021].

HydroBudget (HB) is a spatially distributed groundwater recharge (GWR) model that computes a superficial water budget on grid cells of regional-scale watersheds with outputs aggregated into monthly time steps and with limited computational time. The model is open-source and was coded in R. This `rechaRge` R package is the result of the effort to make this code reusable and extensible.

This book is organized into xxx parts. In the Preamble, we introduce the book, provides a tutorial for ...

The xxx part, ...

Finally, the Appendix highlights our contributors.

```{r include=FALSE}
## automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
library(kableExtra)
```
